A tool set for random number generation on GPUs in R

### 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3008 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article,nojss]{jss}

\usepackage{float}
\usepackage{framed}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{listings}
\usepackage[shortcuts]{extdash}
\usepackage{booktabs}
\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }

@@

<<preliminaries, echo=FALSE, results="hide">>=
  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
  options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
library("MASS")
@

@@

\begin{leftbar}
% Why?
% 
% gpu massively parallel and relatively cheap, useful for stats
% there are gpu packages in R, gpuR
% no random numbers yet.  complicated to do on gpu, need multiple seeds becuase of parallel generation
% reproducability: need to save current state and restore

@@

\end{leftbar}
<<echo=FALSE, results="hide", message=FALSE>>=
library("gpuR")
library("clrng")
setContext(   grep('gpu', listContexts()$device_type) [1]    )
#setwd("/home/ruoyong/diseasemapping/pkg/clrng/inst/documents/paper1_2021_5")
@

@@

\subsection{Using streams to generate uniform random numbers}
The streams created sequentially are then used by the GPU's work-items to generate random numbers. In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be greater than the number of total work-items in use. The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}. Users can set \code{verbose=2} in \fct{runif} to print out the kernels. Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}. Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments. \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively. \code{Npad} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here). \code{index} represents the position of a work-item when mapping it to a one-dimensional data structure. Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to the work-item. The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type. If to generate random integers, \code{temp} is not scaled. At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
\begin{framed}
\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
__kernel void mrg31k3pMatrix(
  __global int* streams,
  __global float* out,
  int Nrow, int Ncol, int Npad){

@@

}//kernel
\end{lstlisting}
\end{framed}
%Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example. Suppose we have a $2 \times 4$ NDRange ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers. In each grid of (a) that represents a work-item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the 2-dimensional global ID of the work-item. Each stream object is passed to each work-item. The number in a cell of the matrix corresponds to the stream index, it indicates which stream generates a random number for this matrix cell. %For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
% \begin{figure}[H]
%     \centering
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}
%     \qquad
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}
%     \caption{Using streams to produce a $4 \times 6$ matrix of random U(0,1) numbers. \label{fig:1}}
% \end{figure}

@@

\begin{algorithm*}[H] \label{algorithm1}
\SetAlgoLined
 1, Generate $U_1$, $U_2$ i.i.d. from $U (0,1)$ \;
 2, Define \begin{align*} 
& R = \sqrt{-2*\log U_1},\\
&  \Theta = 2\pi*U_2,\\
&  X=R*\cos(\Theta),\\
& Y=R*\sin(\Theta);\;\end{align*}
 3, Take $X$ and $Y$ as two independent draws from $N(0,1)$\;
 \caption{Box-Muller algorithm.}
\end{algorithm*}

Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers. Aside from using local memory and the part that does the transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}. If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages. The kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups. \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively. \code{PI_2} is defined to be $\pi/2$. As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
%
\begin{framed}
\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}]
__kernel void mrg31k3pMatrix(
  __global int* streams,
  __global double* out,
  int Nrow, int Ncol, int Npad, int NpadStreams){

@@

      barrier(CLK_LOCAL_MEM_FENCE);
  }//Dcol
}//Drow
streamsFromPrivate(streams,g1,g2,startvalue);
}//kernel
\end{lstlisting}
\end{framed}
%We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange. %Each stream is copied to the private memory of each work-item. 
%Work-items are organized to be in 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method. The shading on the grids in the NDRange distinguishes work-groups. The numbers in the third row in each gird of NDRange represents the local index of each work-item. Each work-group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work-item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work-item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair. %Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously. %Therefore, each work-item generates one Gaussian random number in each iteration, The matrix filling process is the same as that shown in Figure \ref{fig:1}.

% \begin{figure}[H]%
%     \centering
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
%     \qquad
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
%     \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
%     \label{fig3}
% \end{figure}

@@

 <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
library(knitr)
knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), 
  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

@@

\subsection{Comparing run-time: Week data example}\label{fisher_week}
Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
library(latex2exp)
dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
week<-as.matrix(dataweek[,-1])
rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center", latex_options = "HOLD_position")
@

@@

\subsection{A summary of the results}
We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE>>=
library(kableExtra)
dt <- data.frame(
  var1 = c("B", "1M", "10M", "1M", "10M"),
  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 10.74,  63.24),
  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 15.00,  91.58),
  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.72,   4.18),
  var6 = c('Data','month','week','month','week')
)

knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%
  kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
  add_header_above(c(" " = 1, "Computer 1" = 2, "Computer 2" = 2, " " = 1))
@

\begin{figure}[H]
\centering
<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
abline(v = result_month$threshold, col = "blue", lwd = 1.5)

hist(result_week$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
abline(v = result_week$threshold, col = "blue", lwd = 1.5)
@
\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''. The test statistic values of the observed tables are marked with a blue line on each plot. \label{fig4}}
\end{figure}

@@

\begin{algorithm*}[H] \label{algorithm2}
\SetAlgoLined
 1, Calculate the covariance matrix $\Sigma$ between locations\;
 2, Compute the Cholesky decomposition of $\Sigma = L \cdot D \cdot L^\top$, $L$ is the lower unit triangular matrix, $D$ is a diagonal matrix\;
 3, Generate on GPU a random matrix $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix\;
 4, Compute the random samples from $U$ in batches using $U = L \cdot D^{\frac{1}{2}} \cdot Z$ \;
 \caption{Gaussian random fields simulation using covariance matrix decomposition method.}
\end{algorithm*}

@@

\subsection{Simulating Gaussian random fields with Mat\'ern covariances}
<<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=
library("gpuR")
library("clrng")
library("gpuBatchMatrix")
library('geostatsp')
@
Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
<<setupcoords, eval=TRUE>>=
Ngrid = 80
NlocalCache = 1000
Nglobal = c(128, 64, 2)
Nlocal = c(4, 2, 2)
theType = "double"
@

>>>>

\documentclass[article]{jss}

<<downloadJssClassFile, eval=FALSE, include=FALSE>>=
zUrl = 'http://www.jstatsoft.org/public/journals/1/jss-style.zip'
zFile = basename(zUrl)
if(!file.exists(zFile))
        download.file(zUrl, zFile)
unzip(zFile, exdir='.',
                files = grep("^jss", unzip(zFile, list=TRUE)$Name, value=TRUE)
)
@

%\usepackage{float}
%\usepackage{framed}
\usepackage{subcaption}
\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{listings}
\usepackage[shortcuts]{extdash}
\usepackage{booktabs}

@@

<<setup, cache=FALSE,include=FALSE>>=
library('knitr')

options(continue="+  ",  prompt="R> ", digits=3, width=57)

opts_chunk$set(echo=FALSE,fig.path='Figures/G', fig.height=3,
                fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',
                tidy.opts=list(blank=FALSE, width.cutoff=57),
                prompt=TRUE, highlight=FALSE,cache.stuff=1)

hook_output <- function(x, options) {
        if (knitr:::output_asis(x, options)) return(x)
        paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
}

knit_hooks$set(output = hook_output)
knit_hooks$set(message = hook_output)
knit_hooks$set(warning = hook_output)

knit_hooks$set(source  =function(x, options) {
        x = gsub("\n", "\n+", x)
        x = paste("R>",paste(x, collapse='\nR> '))
#       paste0(c('\\begin{example*}', x, '\\end{example*}', ''),
#                       collapse = '\n')
x =  paste0(c('\\begin{CodeInput}', x, '\\end{CodeInput}', ''),
                        collapse = '\n')
                        x               
})

knit_hooks$set(plot = function (x, options) 
                {
                        paste( hook_plot_tex(x, options), "\n", 
                                        sep = "")
                }
)

knit_hooks$set(chunk  = function(x, options) {
        if (knitr:::output_asis(x, options)) return(x)
        theend = gregexpr("end\\{Code(Input|Output)\\}", x)
        theend=theend[[1]]
        if(theend[1]>0){
                theend = theend[length(theend)] + attributes(theend)$match.length[length(theend)]
                x = paste(substr(x,1,theend-1),"\n\\end{CodeChunk}", substr(x,theend, nchar(x)))
                x=sub('begin\\{Code', 'begin{CodeChunk}\n\\\\begin{Code', x)
        }
        x
})
knitr::knit_hooks$set(margins = function(before, options, envir) {
    if (!before) 
        return()
    graphics::par(mar = c(1.5 + 0.9 * options$margins, 1.5 + 
        0.9 * options$margins, 0.2, 0.2), mgp = c(1.45, 0.45, 
        0), cex = 1.25)
})
opts_knit$set(header='')

@

<<preliminaries, echo=FALSE, results="hide", eval=FALSE>>=
  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
  options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF00', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
library("MASS")
@

@@

% Why?
% 
% gpu massively parallel and relatively cheap, useful for stats
% there are gpu packages in R, gpuR
% no random numbers yet.  complicated to do on gpu, need multiple seeds becuase of parallel generation
% reproducability: need to save current state and restore

@@

<<echo=FALSE, results="hide", message=FALSE>>=
library("gpuR")
library("clrng")
setContext(   grep('gpu', listContexts()$device_type) [1]    )
#setwd("/home/ruoyong/diseasemapping/pkg/clrng/inst/documents/paper1_2021_5")
@

@@

\subsection{Using streams to generate uniform random numbers}
The streams created sequentially are then used by the GPU's work-items to generate random numbers. In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be greater than the number of total work-items in use. The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}. Users can set \code{verbose=2} in \fct{runif} to print out the kernels. Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}. Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments. \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively. \code{Npad} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here). \code{index} represents the position of a work-item when mapping it to a one-dimensional data structure. Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to the work-item. The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type. If to generate random integers, \code{temp} is not scaled. At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.

\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
__kernel void mrg31k3pMatrix(
  __global int* streams,
  __global float* out,
  int Nrow, int Ncol, int Npad){

@@

}//kernel
\end{lstlisting}

%Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example. Suppose we have a $2 \times 4$ NDRange ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers. In each grid of (a) that represents a work-item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the 2-dimensional global ID of the work-item. Each stream object is passed to each work-item. The number in a cell of the matrix corresponds to the stream index, it indicates which stream generates a random number for this matrix cell. %For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
% \begin{figure}[ht]
%     \centering
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}
%     \qquad
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}
%     \caption{Using streams to produce a $4 \times 6$ matrix of random U(0,1) numbers. \label{fig:1}}
% \end{figure}

@@

\begin{algorithm*}[ht] \label{algorithm1}
\SetAlgoLined
 1, Generate $U_1$, $U_2$ i.i.d. from $U (0,1)$ \;
 2, Define \begin{align*} 
& R = \sqrt{-2*\log U_1},\\
&  \Theta = 2\pi*U_2,\\
&  X=R*\cos(\Theta),\\
& Y=R*\sin(\Theta);\;\end{align*}
 3, Take $X$ and $Y$ as two independent draws from $N(0,1)$\;
 \caption{Box-Muller algorithm.}
\end{algorithm*}

Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers. Aside from using local memory and the part that does the transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}. If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages. The kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups. \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively. \code{PI_2} is defined to be $\pi/2$. As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
%

\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}]
__kernel void mrg31k3pMatrix(
  __global int* streams,
  __global double* out,
  int Nrow, int Ncol, int Npad, int NpadStreams){

@@

      barrier(CLK_LOCAL_MEM_FENCE);
  }//Dcol
}//Drow
streamsFromPrivate(streams,g1,g2,startvalue);
}//kernel
\end{lstlisting}

%We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange. %Each stream is copied to the private memory of each work-item. 
%Work-items are organized to be in 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method. The shading on the grids in the NDRange distinguishes work-groups. The numbers in the third row in each gird of NDRange represents the local index of each work-item. Each work-group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work-item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work-item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair. %Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously. %Therefore, each work-item generates one Gaussian random number in each iteration, The matrix filling process is the same as that shown in Figure \ref{fig:1}.

% \begin{figure}[ht]%
%     \centering
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
%     \qquad
%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
%     \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
%     \label{fig3}
% \end{figure}

@@

 <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h'>>=
library(knitr)
knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), 
  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
@

@@

\subsection{Comparing run-time: Week data example}\label{fisher_week}
Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
library(latex2exp)
dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
week<-as.matrix(dataweek[,-1])
rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
@

@@

\subsection{A summary of the results}
We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE, fig.pos='h'>>=
library(kableExtra)
dt <- data.frame(
  var1 = c("B", "1M", "10M", "1M", "10M"),
  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 10.74,  63.24),
  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 15.00,  91.58),
  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.72,   4.18),
  var6 = c('Data','month','week','month','week')
)

knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
  add_header_above(c(" " = 1, "Computer 1" = 2, "Computer 2" = 2, " " = 1))
@

<<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "Approximate sampling distributions of the test statistics from the two examples Month and Week. The test statistic values of the observed tables are marked with a blue line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1>>=
hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),
     breaks=40, ylab="proportion", main="", prob=TRUE)
abline(v = result_month$threshold, col = "blue", lwd = 1.5)

hist(result_week$cpu, xlab='test statistic', ylab="proportion", prob=TRUE, main="")
abline(v = result_week$threshold, col = "blue", lwd = 1.5)
@

@@

\begin{algorithm*}[ht] \label{algorithm2}
\SetAlgoLined
 1, Calculate the covariance matrix $\Sigma$ between locations\;
 2, Compute the Cholesky decomposition of $\Sigma = L \cdot D \cdot L^\top$, $L$ is the lower unit triangular matrix, $D$ is a diagonal matrix\;
 3, Generate on GPU a random matrix $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix\;
 4, Compute the random samples from $U$ in batches using $U = L \cdot D^{\frac{1}{2}} \cdot Z$ \;
 \caption{Gaussian random fields simulation using covariance matrix decomposition method.}
\end{algorithm*}

@@

\subsection{Simulating Gaussian random fields with Mat\'ern covariances}
<<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=
library("gpuR")
library("clrng")
library("gpuBatchMatrix")
library('geostatsp')
@
Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
<<setupcoords, eval=TRUE>>=
NlocalCache = 1000
Nglobal = c(128, 64, 2)
Nlocal = c(4, 2, 2)
theType = "double"
@

<<swissRainBoundary, echo=TRUE>>=
data("swissRain", package="geostatsp")
myRaster = geostatsp::squareRaster(swissBorder, 60)
myRaster
@
Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets. Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuBatchMatrix::maternGpuParam}.
<<setup paramsBatch, eval=TRUE,echo=FALSE>>=
params = 
rbind(c(shape=1.25, range=50*1000, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
c(shape=2.15, range=60*1000, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
c(shape=0.6, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7),
c(shape=3, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7)
)
#c(shape=2.15, range=40*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
@
<<show parameterBatch,eval=TRUE>>=
params
#paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
@
Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices 
are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
<<r outputBatchCreate, eval=TRUE>>=
#maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
#                         nrow(coordsGpu),type=theType)
maternCov = gpuBatchMatrix::maternBatch(
  params, myRaster,          
  Nglobal=c(128,64), Nlocal=c(16,4))
dim(maternCov)
@
Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition. Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}. Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
<<r cholBach, eval=TRUE>>=
diagMat = gpuBatchMatrix::cholBatch(maternCov, 
                          Nglobal= c(128, 8), Nlocal= c(32, 8))
@
\begin{gather}
\begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
\end{bmatrix}
 \rightarrow
 \begin{bmatrix}
  L_{1} \\ L_{2} \\L_{3} \\ \vdots
  \end{bmatrix} \text{and}
  \begin{bmatrix}
  D_{1} \\ D_{2} \\D_{3} \\ \vdots
  \end{bmatrix}
\end{gather}
Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
<<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
streamsGpu <- createStreamsGpu(n=128*64)
@
<<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
                 type = theType)
@
Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration. \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide'>>=
simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
                   type = gpuR::typeof(zmatGpu))
